---
title: "EF_feather_plot.Rmd"
author: "Brooke Sevchik"
date: "2025-09-24"
output: html_document
---

# Pre-Plotting

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(forcats)
library(ssh)
```

```{r load in necessary files}
# Load in acq_time_data and input data from cubic - requires you to mount cubic onto laptop for this to run

input_data <- read.csv('/Volumes/cbica/projects/executive_function/code/data/input_data_feather_plot.csv')

acq_time_data <- read.csv('/Volumes/cbica/projects/executive_function/data/bids/EF_bids_data/all_session_times.csv')

scanned_participants <- read.csv('/Users/bsevchik/projects/EF_clinical/participants.csv')

# Load in subject data from remote server (will prompt password)
session <- ssh_connect("bsevchik@bblrepo1.pmacs.upenn.edu")  # will prompt once for password
csv_text <- ssh_exec_internal(session, "cat /data/secure/bit_dwh/oracle/subject.csv")
subject_data <- read.csv(text = rawToChar(csv_text$stdout))
ssh_disconnect(session)


```



Get DOB for participants in EF
```{r dob}
# just select columns you need
subject_data <- subject_data %>%
  select(BBLID, DOBIRTH)

# Strip "sub-" from scanned_participants IDs
scanned_ids <- scanned_participants %>%
  mutate(BBLID = sub("^sub-", "", participant_id)) %>%
  pull(BBLID)

# Filter subject_data to only keep matching IDs
subject_data <- subject_data %>%
  filter(BBLID %in% scanned_ids)
subject_data
```


Add in columns 'session-id' (all filled in with value '1') and acq_time (leave blank for time)
```{r add new columns}
# Add new columns
input_data$session_id <- "1"
input_data$acq_time   <- ""            # leave blank for now
input_data$dob        <- ""  #leave blank for now

```

Now, using acq_time_data, fill in subsequent sessions for each participant, and fill in acquisition time for each session.
```{r load in necessary files}

library(dplyr)
library(tidyr)

# Ensure input_data has consistent types
input_data <- input_data %>%
  mutate(
    participant_id = as.character(participant_id),  # e.g., "1"
    session_id     = as.character(session_id),      # currently "1" for all
    acq_time       = as.character(acq_time)
  )

# Normalize acq_time_data: strip "sub-" and "ses-"
acq_norm <- acq_time_data %>%
  mutate(
    subject          = sub("^sub-", "", as.character(subject)),  # "sub-1" -> "1"
    session          = as.character(session),
    acquisition_time = as.character(acquisition_time),
    session_id       = sub("^ses-", "", session)                 # "ses-1" -> "1"
  ) %>%
  filter(session_id %in% c("1","2","3"))

# --- 1) Update acq_time for session 1 in input_data
acq_s1 <- acq_norm %>%
  filter(session_id == "1") %>%
  select(participant_id = subject, session_id, acq_time = acquisition_time)

input_s1_updated <- input_data %>%
  left_join(acq_s1, by = c("participant_id","session_id")) %>%
  mutate(acq_time = dplyr::coalesce(acq_time.y, acq_time.x)) %>%
  select(-acq_time.x, -acq_time.y)

# --- 2) Add rows for sessions 2–3, leaving other columns blank
acq_s23 <- acq_norm %>%
  filter(session_id %in% c("2","3")) %>%
  transmute(participant_id = subject, session_id, acq_time = acquisition_time)

all_cols   <- names(input_s1_updated)
blank_cols <- setdiff(all_cols, c("participant_id","session_id","acq_time"))

if (nrow(acq_s23) == 0) {
  # No ses-2/3 rows exist; create a 0-row skeleton
  new_rows <- input_s1_updated[0, all_cols]
} else {
  new_rows <- acq_s23
  
  # Decide which blanks should be "" vs NA based on target column types
  is_char <- sapply(input_s1_updated[blank_cols], function(x) is.character(x) || is.factor(x))
  is_num  <- sapply(input_s1_updated[blank_cols], is.numeric)
  
  for (cn in names(is_char)[is_char]) new_rows[[cn]] <- rep_len("", nrow(new_rows))
  for (cn in names(is_num)[is_num])   new_rows[[cn]] <- rep(NA_real_, nrow(new_rows))
  
  other_cols <- setdiff(blank_cols, c(names(is_char)[is_char], names(is_num)[is_num]))
  for (cn in other_cols) new_rows[[cn]] <- rep(NA_character_, nrow(new_rows))
  
  new_rows <- new_rows %>% select(all_of(all_cols))
}

# --- 3) Combine and tidy
updated_input <- bind_rows(input_s1_updated, new_rows) %>%
  arrange(participant_id, suppressWarnings(as.integer(session_id)))


```

Fill in DOB for each participant
```{r add DOB}
# Ensure join keys are the same type
subject_data <- subject_data %>% mutate(BBLID = as.character(BBLID))
updated_input <- updated_input %>% mutate(participant_id = as.character(participant_id))

# Join DOBIRTH onto updated_input and overwrite dob
updated_input <- updated_input %>%
  left_join(subject_data %>% select(BBLID, DOBIRTH),
            by = c("participant_id" = "BBLID")) %>%
  mutate(dob = dplyr::coalesce(DOBIRTH, dob)) %>%
  select(-DOBIRTH)
```

Now, fill in study_group, sex, race, and ethnicity blanks for session 2 and 3 with the same info as from session 1. 
```{r fill in demographics info}
demog_cols <- c("study_group","sex","race","ethnicity")

# ensure character so "" checks work
updated_input <- updated_input %>%
  mutate(across(any_of(demog_cols), ~as.character(.)))

# build lookup from session 1
demog_lookup <- updated_input %>%
  filter(session_id == "1") %>%
  select(participant_id, dplyr::all_of(demog_cols)) %>%
  distinct()

# join ses-1 values
updated_input <- updated_input %>%
  left_join(demog_lookup, by = "participant_id", suffix = c("", ".s1"))

# fill only for sessions 2/3 where blank/NA
fill_mask <- updated_input$session_id %in% c("2","3")
for (cn in demog_cols) {
  s1_col <- paste0(cn, ".s1")
  # if target col is missing/blank and it's session 2/3, copy from s1
  to_fill <- fill_mask & (is.na(updated_input[[cn]]) | updated_input[[cn]] == "")
  updated_input[[cn]][to_fill] <- updated_input[[s1_col]][to_fill]
}

# clean up helper columns
updated_input <- updated_input %>% select(-dplyr::ends_with(".s1"))


```


Now, calculate age (in months & years) for sessions 2 and 3, using info from acq_time for all sessions and DOB
```{r calculate age}

# Start with the combined data
df <- if (exists("updated_input")) updated_input else input_data

df <- df %>%
  mutate(
    # Ensure strings
    dob_chr = as.character(dob),
    acq_chr = as.character(acq_time),

    # Parse DOB (support "YYYY-MM-DD" and "MM/DD/YYYY")
    dob_date = dplyr::coalesce(
      suppressWarnings(ymd(dob_chr, quiet = TRUE)),
      suppressWarnings(mdy(dob_chr, quiet = TRUE))
    ),

    # Parse acquisition time (supports "YYYY-MM-DDTHH:MM:SS[.ffffff]")
    acq_dt = suppressWarnings(ymd_hms(acq_chr, quiet = TRUE, tz = "UTC")),
    acq_date = dplyr::coalesce(as_date(acq_dt),
                               suppressWarnings(ymd(acq_chr, quiet = TRUE)))
  ) %>%
  rowwise() %>%
  mutate(
    age_months = if (is.na(dob_date) || is.na(acq_date))
      NA_integer_ else as.integer(floor(time_length(interval(dob_date, acq_date), "months"))),
    age = if (is.na(dob_date) || is.na(acq_date))
      NA_real_ else round(time_length(interval(dob_date, acq_date), "years"), 2)
  ) %>%
  ungroup() %>%
  select(-dob_chr, -acq_chr, -acq_dt)  # keep dob_date & acq_date if you find them useful

updated_input <- df


# Get average age across essions
mean_age <- mean(updated_input$age)
print(mean_age)

```








# Plotting

```{r feather plot code}

df <- updated_input %>%
  mutate(
    session_id = factor(session_id, levels = c("1","2","3")),
    age = as.numeric(age)
  ) %>%
  filter(!is.na(age))  # need age to place points/lines

# Order participants by baseline (session 1) age; if missing, fall back to min age
id_order <- df %>%
  group_by(participant_id) %>%
  summarize(
    baseline_age = suppressWarnings(min(age[session_id == "1"], na.rm = TRUE)),
    any_age      = suppressWarnings(min(age, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(baseline_age = ifelse(is.infinite(baseline_age), any_age, baseline_age)) %>%
  arrange(baseline_age) %>%
  mutate(y_index = row_number())

# attach y-position to each row
df_plot <- df %>% left_join(id_order, by = "participant_id")

# one segment per participant spanning min→max observed age
seg_df <- df_plot %>%
  group_by(participant_id, y_index) %>%
  summarize(xmin = min(age, na.rm = TRUE),
            xmax = max(age, na.rm = TRUE),
            .groups = "drop") %>%
  filter(is.finite(xmin), is.finite(xmax))

# ——— Feather plot (segments + colored session points) ———
p <- ggplot() +
  geom_segment(data = seg_df,
               aes(x = xmin, xend = xmax, y = y_index, yend = y_index),
               linewidth = 0.7, color = "grey80") +
  geom_point(data = df_plot,
             aes(x = age, y = y_index, color = session_id),
             size = 2.2) +
  scale_y_continuous(breaks = id_order$y_index, labels = id_order$participant_id) +
  labs(x = "Age (years)", y = "Participant", color = "Session") +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.y  = element_blank(),
    axis.ticks.y = element_blank()
  )

p

ggsave("/Users/bsevchik/projects/EF_clinical/feather_plot/feather_plot.png", p,
       width = 8, height = 12, dpi = 300)  # make height bigger

```


