---
title: "get_nback_sessions_included"
author: "Brooke Sevchik"
date: "2026-05-21"
output: html_document
---

---
title: "subjects_task_contrast"
author: "Brooke Sevchik"
date: "2025-11-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, include=FALSE}
library(tidyverse)
library(dplyr)
library(readr)
library(stringr)
```

```{r load task contrast QC csv}
task_contrast_QC <- read.csv('/Users/bsevchik/Documents/GitHub/transdiagnostic_executive_function/QC/qc_csvs/final_QC_csvs/EF_desc-taskcontrast_qc.csv')

#filter to only include those that passed
task_contrast_passed <- task_contrast_QC %>%
  filter(qc_determination == "pass")
```

Load in EF_age_info.csv, filter that csv to exclude the session IDs not in fmri_data_nback_filtered

```{r filter age info}
#load in age info
EF_age_info <- read_csv("/Users/bsevchik/projects/EF_code/EF_self_report/EF_age_info.csv")
EF_age_info

#filter
EF_age_info_filtered <- EF_age_info %>%
  semi_join(
    task_contrast_passed %>%
      distinct(participant_id, session_id),
    by = c("participant_id", "session_id")
  )
EF_age_info_filtered

#check how many rows filtered
nrow(EF_age_info)
nrow(task_contrast_passed)
nrow(EF_age_info_filtered)
```

#Average demoagrphics (from 105 participants included in task contrast data for neurovault)

Within-subject average from EF_age_info_filtered --> EF_age_info_filtered_avg (one row per subject)
```{r within-subject average}
EF_age_info_subject_avg <- EF_age_info_filtered %>%
  group_by(participant_id) %>%
  summarise(
    age_avg = mean(age, na.rm = TRUE),
    .groups = "drop"
  )
EF_age_info_subject_avg
```


Between-subject mean age, min, and max from EF_age_info_filtered_avg
```{r between-subject average}
EF_age_info_subject_avg %>%
  summarise(
    mean_age = mean(age_avg, na.rm = TRUE),
    min_age = min(age_avg, na.rm = TRUE),
    max_age = max(age_avg, na.rm = TRUE)
  )
```


From one row per subject EF_age_info_filtered_avg, calculate proportion of male subjects
Load in demographics info, then filter.

```{r filter demographic info}
#load in demographic info
EF_demographic_info <- read_tsv("/Users/bsevchik/Documents/GitHub/transdiagnostic_executive_function/demographics/participants.tsv")
EF_demographic_info

EF_demographic_info <- EF_demographic_info %>%
  mutate(
    participant_id = str_remove(participant_id, "sub-"),
    participant_id = as.integer(participant_id)
  )
EF_demographic_info

#filter
EF_demographic_info_filtered <- EF_demographic_info %>%
  semi_join(
    task_contrast_passed %>%
      distinct(participant_id),
    by = c("participant_id")
  )
EF_demographic_info_filtered

#check how many rows filtered
nrow(EF_demographic_info_filtered) #should be 105
```

Calculate proportion of males.

```{r male proportion}
EF_demographic_info_filtered %>%
  summarise(
    n_total = n(),
    n_male = sum(sex == "M", na.rm = TRUE),
    prop_male = mean(sex == "M", na.rm = TRUE)
  )
```


Calculate number of rejected subjects: number of rejected from no nback data/did not pass QC, and then number of rejected from no events.tsv or response_time is all n/a
```{r calculate rejected subjects}
EF_fmri_QC <- read.csv('/Users/bsevchik/Documents/GitHub/transdiagnostic_executive_function/QC/qc_csvs/final_QC_csvs/EF_desc-fMRI_qc.csv')

EF_fmri_QC_nback <- EF_fmri_QC %>%
  filter(task == 'nback')

EF_fmri_QC_nback %>%
  summarise(n_unique_subjects_fmri = n_distinct(participant_id))


task_contrast_QC_filtered <- task_contrast_QC %>%
  filter(qc_determination == "pass")
task_contrast_QC_filtered

task_contrast_QC_filtered %>%
  summarise(n_unique_subjects = n_distinct(participant_id))

#123-105 = 18


```

